### 2017-11-26 notes_jc.tex
implemented model based on MC

\paragraph{A Monte Carlo based method} We repeatedly sample $\tilde Z$ \emph{uniformly} from $\bm Z$ and use the sample mean to approximate summations over $\bm Z$. \begin{align*}
&\log E_Z\pr{p(T\uppr{k} | Z\uppr{k}, \bm W)| \bm W} \\&= \log\bk{ |\bm Z|E_{\tilde Z}\bk{p(T\uppr{k} | \tilde Z\uppr{k},\bm W)p(\tilde Z\uppr{k} | \bm W)} \tag{where $\tilde Z$ is uniform over $\bm Z$}} \\
&= \log E_{\tilde Z}\bk{p(T\uppr{k} | \tilde Z\uppr{k},\bm W)p(\tilde Z\uppr{k} | \bm W)} + \text{const.} \\
&\ge E_{\tilde Z}\bk{\log p(T\uppr{k} | \tilde Z\uppr{k},\bm W)} + E_{\tilde Z}\bk{\log p(\tilde Z\uppr{k} | \bm W)} \tag{Jensen; omitting constant}\\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L \bk{-\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \log \sum_{\bm Z} \exp\pr{-\sum_{Z\uppr{k}} w_i}}  \tag{We parameterize the distribution of $Z\uppr{k}$ as Categorical with $\frac{\exp(-\sum w)}{\sum \exp(-\sum w)}$} \\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L \sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \sum_{i=1}^L \log\bk{|\bm Z| E_{\tilde Z}\bk{\exp\pr{-\sum_{\tilde Z} w_i}}} \\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \sum_{i=1}^L \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i}
\end{align*}

>>>>

\paragraph{A Monte Carlo based method} We repeatedly sample $\tilde Z$ \emph{uniformly} from $\bm Z$ and use the sample mean to approximate summations over $\bm Z$. \begin{align*}
&\log E_Z\pr{p(T\uppr{k} | Z\uppr{k}, \bm W)| \bm W} \\&= \log\bk{ |\bm Z|E_{\tilde Z}\bk{p(T\uppr{k} | \tilde Z\uppr{k},\bm W)p(\tilde Z\uppr{k} | \bm W)} \tag{where $\tilde Z$ is uniform over $\bm Z$}} \\
&= \log E_{\tilde Z}\bk{p(T\uppr{k} | \tilde Z\uppr{k},\bm W)p(\tilde Z\uppr{k} | \bm W)} + \text{const.} \\
&\ge E_{\tilde Z}\bk{\log p(T\uppr{k} | \tilde Z\uppr{k},\bm W)} + E_{\tilde Z}\bk{\log p(\tilde Z\uppr{k} | \bm W)} \tag{Jensen; omitting constant}\\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L \bk{-\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \log \sum_{\bm Z} \exp\pr{-\sum_{Z\uppr{k}} w_i}}  \tag{We parameterize the distribution of $Z\uppr{k}$ as Categorical with $\frac{\exp(-\sum w)}{\sum \exp(-\sum w)}$} \\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L \sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \frac{1}{L}\sum_{i=1}^L \log\bk{|\bm Z| E_{\tilde Z}\bk{\exp\pr{-\sum_{\tilde Z} w_i}}} \\
&= -\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} - \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 - \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i - \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i}
\end{align*}
